\section{Quadratic: degree=2}
\label{sec.quadratic}

\input{pfg-parabola}

A quadratic polynomial is of the form $P(x)=a x^2 + b x + c$.  If
$a=0$, then $P(x)$ is actually a first degree polynomial which can be
solved using the techniques described in Section~\ref{sec.line}.

We have the plots of three quadratic polynomials, $P_1(x)$, $P_2(x)$,
and $P_3(x)$, shown in Figure~\ref{fig.parabola}.  Each plot takes the
geometric form of a parabola.  As can be seen in the figure, the
parabola might intersect the x-axis twice (\eg, $P_1(x)$), once (\eg,
$P_2(x)$, or not at all (\eg, $P_3(x)$).

\begin{align*}
  P_1(x) &= 2x^2 - 2 x - 5\\
  P_2(x) &= x^2 -x + \frac{1}{4}\\
  P_3(x) &=  -2x^2 + 2x -2
\end{align*}


\subsection{Mathematically Computing the Roots of a Quadratic}

To determine how many times a parabola touches the x-axis, we need to
find the roots of the corresponding quadratic (degree 2) polynomial of
the form $a x^2 + b x + c$.  The formula is called the \emph{quadratic
formula} is given in Equation~\eqref{eq.quad.formula}.  In particular,
we need to look at the expression under the square root, called the
discriminant: $b^2 - 4a c$.
\begin{itemize}
\item If the discriminant is positive, then the polynomial has two
  distinct real roots.
\item If the discriminant is zero, then the positive and negative
  square roots are both 0; thus the polynomial has 1 single root.
\item If the discriminant is negative, then there is a negative under
  the radical sign; thus the polynomial has no real roots.  The
  polynomial does have complex roots; however we will ignore complex
  roots for this atelier.
\end{itemize}

Example~\ref{ex.degree2.p1} shows the computation of the two distinct roots of $P_1(x)$.

\begin{example}{Explicit Computation of Roots of $P_1(x)$}{degree2.p1}
  $P_1(x)$, has two roots.  They can be found using the quadratic formula.
  \begin{align*}
    a&=2\\
    b&=-2\\
    c&=-5\\
    b^2 - 4a c &= (-2)^2 - 4 (2) (-5)\\
    &= 4 + 40 = 44\\
    \sqrt{b^2 - 4a c} &= \sqrt{44} = 2\sqrt{11}
  \end{align*}

\begin{align}
  x &= \frac{-b~ \pm \sqrt{b^2 - 4a c}}{2a}\label{eq.quad.formula}\\
  &= \frac{-(-2)~ \pm 2\sqrt{11}}{(2)(2)}\nonumber\\
  &= \frac{2 ~\pm 2\sqrt{11}}{4}
  = \frac{1}{2} \pm \frac{1}{2}\sqrt{11}\nonumber
\end{align}
  
\end{example}

Example~\ref{ex.degree2.p2} shows the computation of the root of $P_2(x)$.

\begin{example}{Explicit Computation of Root of $P_2(x)$}{degree2.p2}
In the case of $P_2(x) = x^2 -x + \frac{1}{4}$, we do a similar computation.

\begin{align*}
  a &= 1\\
  b &= -1\\
  c &= \frac{1}{4} \\
  b^2 - 4a c &= (-1)^2 - 4 (1) (\frac{1}{4})\\
  &= 1 - 1 = 0
\end{align*}

\begin{align*}
  x &= \frac{-b~ \pm \sqrt{b^2 - 4a c}}{2a}\\
  &= \frac{-(-1)~ \pm \sqrt{0}}{(2)(1)}\\
  &= \frac{1}{2}
\end{align*}


Thus there is exactly one root.  The existance of the single root can be
seen in Figure~\ref{fig.parabola} by the fact that the
parobala touches the x-axis tangentially.
\end{example}

Example~\ref{ex.degree2.p3} examines $P_3(x)$.

\begin{example}{Examination of $P_3(x)$}{degree2.p3}
  In the case of $P_3(x) = -2x^2 + 2x -2$, we may attempt the same but we have a negative discriminant.
  \begin{align*}
    a&=-2\\
    b&=2\\
    c&=-2\\
    b^2 - 4a c  &=(-2)^2 - (4)(-2)(-2)\\
    &= 4 - 16 = -12 < 0
  \end{align*}

When the discriminant
is negative, there are no \emph{real} roots.  This can be seen in
Figure~\ref{fig.parabola} by the fact that the parobola does not
touch the x-axis at all.
\end{example}

\subsection{Programmatically Computing Roots of a Quadratic}

Steps for computing roots of a polynomial of degree 2.
\begin{enumerate}
\item Assume the coefficients of $P(x)$ are in the variables \code{a}, \code{b}, and \code{c}.
\item If \code{a == 0}, then delegate to previous solution by calling the function \code{find\_x\_intercept} and returning its value.  Note that \code{find\_x\_intercept} takes two arguments, which it calls \code{a} and \code{b}, but these are not \code{a} and \code{b} of the quadratic.   In fact for \code{find\_x\_intercept}, \code{a} is the coefficients of $x$ and \code{b} is the bias term of $a x + b$.  However, for the quadratic, \code{a} is the coefficient of $x^2$, and \code{b} is the code of $x$ in $a x^2 + b x + c$.
\item Compute the discriminant, storing in the variable \code{discriminant}.
\item Test three conditions.
  \begin{enumerate}
  \item If the discriminant is positive, then return a list of the two roots.
  \item If the discriminant is zero (or very close to zero), then return a list (of length two) of the same root twice.
  \item If the discriminant is negative ( $< -\varepsilon$), then return an empty list~\code{[]}.
  \end{enumerate}
\end{enumerate}


\begin{listing}{Function to compute roots of a quadratic polynomial.}{code.quadratic}
\begin{minipage}[c]{0.95\textwidth}\begin{lstlisting}
def find_quadratic_roots(a, b, c):
    epsilon = 0.001
    discriminant = b * b - 4 * a * c
    if a == 0:
        # invoke the previous solution for x-intercept of a line
        # CHALLENGE: student must complete the implementation.
        # HINT: goal = 1 line
        raise NotImplementedError()

    if abs(discriminant) < epsilon:
        # Here we know the discriminant is very close to zero,
        #   so return a singleton list containing the one root.
        # CHALLENGE: student must complete the implementation.
        # HINT: goal = 1 line
        raise NotImplementedError()

    elif discriminant > 0:
        # Here we know the discriminant is not close to zero,
        #   and we also know it is positive, so we return a
        #   list containing the two roots according to the
        #   quadratic formula.
        # CHALLENGE: student must complete the implementation.
        # HINT: goal <= 2 lines
        raise NotImplementedError()

    else:
        # Here we know the discriminant is negative, so return an
        #    empty list of no roots.
        # CHALLENGE: student must complete the implementation.
        # HINT: goal = 1 line
        raise NotImplementedError()

\end{lstlisting}\end{minipage}\end{listing}

